\textcolor[rgb]{0,0,1}{Data-Driven Model Predictive Control for Energy and Thermal Comfort Optimization in Building Energy Systems using Random Forests}

### 2018-01-20 paper/final/casestudyEPlus.tex
small changes

In the following simulations, we consider a long DR event from 7am - 2pm when the end-users are expected to follow/track the reference power signal sent by the utility. This is indeed common in Demand Tracking Control. 
During offline training, we sample data every 15 min to learn 2 kinds of forests. (1) Power forests are built using output as the total building power consumption, and (2) Temperature forests with output as temperature of one of the 22 zones. 
The training data set contains the following types of features. (1) The \textit{weather data} which includes measurements of the outside air temperature and relative humidity. Since we are interested in predicting the power consumption or the zone temperature for a finite horizon, we include the weather forecast of the complete horizon in the training features. (2) The \textit{schedule data} includes the proxy variables which correlate with repeated patterns of electricity consumption e.g., due to occupancy or equipment schedules. Day of Week is a categorical predictor which takes values from 1-7 depending on the day of the week. This variable can capture any power consumption patterns which occur on specific days of the week. Likewise, Time of Day is quite an important predictor of power consumption as it can adequately capture daily patterns of occupancy, lighting and appliance use without directly measuring any one of them. Besides using proxy schedule predictors, actual building equipment schedules can also be used as training data for building the trees. (3) The \textit{building data} include (i) cooling set points for  the guest rooms, kitchen and corridors, (ii) supply air temperature, and (iii) chilled water temperature.
For the following simulations, we use five control variables (i) cooling set point for corridors $\mathsf{ClgSP}$, (ii) cooling set point for guest rooms $\mathsf{GuestSP}$, (iii) cooling set point for kitchen $\mathsf{KitchenSP}$, (iv) chilled water supply temperature $\mathsf{ChwSP}$, and (v) supply air temperature  $\mathsf{SupplyAirSP}$, so $\tX^c = [\mathsf{ClgSP},\mathsf{GuestClgSP},\mathsf{KitchenClgSP},\mathsf{SupplyAirSP}, \mathsf{ChwSP}]$. 
The power forest $\mathcal{R}_p$ is built using the total building power consumption $\tP$. Its features $\tX^d$ include the weather variables, their lag terms and their forecast over the horizon, the schedule variables, and finally the lag terms of the power consumption.
The temperature forest $\mathcal{R}_t$ is built with zone temperature $\tT$ as the output. Except for the lag terms corresponding to the same zone temperature, all other features are same in $\tX^d$.
Figure~\ref{F:sepvars} shows the prediction accuracy for the power forest, and also explains the two level training approach introduced in Section~\ref{SS:sepvar}. During S1, the forests are trained using only disturbances as the features. Then in S2, the local effects of the control variables are accounted for by the linear models in the leaves. We observe how the accuracy is drastically improved after including the linear models in the predictions.

\subsection{Power Management}
\label{SS:powermanagement}
Typically, the end customer receives a notification to curtail the power by some fraction. 
In this example on power management, we show how DPC can generate optimal inputs to track a desired power signal within a small allowance while maintaining the zone level thermal comfort. It may not be possible to have the same thermal comfort level in all the zones due to power curtailment, so we choose one zone (for example CEO's office) where the constraints must be met.
\begin{figure}[t!]
	\centering
	\includegraphics[width=30pc]{figures/eplus_validation.eps}	
	\caption{Model accuracy during training: The prediction made by forest using only $\tX^d$ (red) captures the effect due to disturbances. The linear models in the leaves capture the local effects (green) due to the control inputs $\tX^c$ and improve the model accuracy.}
	\label{F:sepvars}
	\captionsetup{justification=centering}
\end{figure}
This is done by solving the following optimization problem with control variables $\tX^c = [\mathsf{ClgSP},\mathsf{GuestClgSP},\mathsf{KitchenClgSP},\mathsf{SupplyAirSP}, \mathsf{ChwSP}]$ as defined before:
\begin{align}
\begin{aligned}
\text{min } & \ \ \ \ \ \ \ \sum_{j=1}^{N} ({\tP}_{\mathrm{k+j|k}} - \tP_{\mathrm{ref}})^2 +  \lambda\epsilon_j + \nu \delta_j \\
\text{s.~t. } & \ \ \ \ \ \tP_{\mathrm{k+j|k}} =  \hat{\Theta}^T_{\tP_j} [ 1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
& \ \ \ \ \ \tT_{\mathrm{k+j|k}} =  \hat{\Theta}^T_{\tT_j} [ 1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tP}-\epsilon_j \leq \tP_{\mathrm{k+j|k}} \leq \bar{\tP} + \epsilon_j\\
& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tT}-\delta_j \leq \tT_{\mathrm{k+j|k}} \leq \bar{\tT} + \delta_j\\
& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tX}^c \leq \tX^c_{\mathrm{k+j-1|k}} \leq \bar{\tX}^c\\ 
& \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0,\delta_j \geq 0, \ j = 1,\dots,N.
\end{aligned}
\label{E:ptrack}
\end{align}

>>>>

In the following simulations, we consider a long DR event from 7am - 2pm when the end-users are expected to follow/track the reference power signal sent by the utility. This is indeed common in Demand Tracking Control. 
During offline training, we sample data every 15 min to learn 2 kinds of forests. (1) Power forests are built using output as the total building power consumption, and (2) Temperature forests with output as temperature of one of the 22 zones. 
The training data set contains the following types of features. (1) The \textit{weather data} which includes measurements of the outside air temperature and relative humidity. Since we are interested in predicting the power consumption or the zone temperature for a finite horizon, we include the weather forecast of the complete horizon in the training features. (2) The \textit{schedule data} includes the proxy variables which correlate with repeated patterns of electricity consumption e.g., due to occupancy or equipment schedules. Day of Week is a categorical predictor which takes values from 1-7 depending on the day of the week. This variable can capture any power consumption patterns which occur on specific days of the week. Likewise, Time of Day is quite an important predictor of power consumption as it can adequately capture daily patterns of occupancy, lighting and appliance use without directly measuring any one of them. Besides using proxy schedule predictors, actual building equipment schedules can also be used as training data for building the trees. (3) The \textit{building data} include (i) cooling set points for  the guest rooms, kitchen and corridors, (ii) supply air temperature, and (iii) chilled water temperature.
For the following simulations, we use five control variables (i) cooling set point for corridors $\mathsf{ClgSP}$, (ii) cooling set point for guest rooms $\mathsf{GuestSP}$, (iii) cooling set point for kitchen $\mathsf{KitchenSP}$, (iv) chilled water supply temperature $\mathsf{ChwSP}$, and (v) supply air temperature  $\mathsf{SupplyAirSP}$, so $u := [\mathsf{ClgSP},\mathsf{GuestClgSP},\mathsf{KitchenClgSP},\mathsf{SupplyAirSP}, \mathsf{ChwSP}]^\top$. 
The power forest $\mathcal{F}_p$ is built using the total building power consumption $\tP$. Its features in $\X^d$ include the weather variables, their lag terms and their forecast over the horizon, the schedule variables, and finally the lag terms of the power consumption.
The temperature forest $\mathcal{R}_t$ is built with zone temperature $\tT$ as the output. Except for the autoregressive terms corresponding to the same zone temperature, all other features are same in $\X^d$.
Figure~\ref{F:sepvars} shows the prediction accuracy for the power forest, and also explains the two level training approach introduced in Section~\ref{SS:sepvar}. During S1, the forests are trained using only disturbances as the features. Then in S2, the local effects of the control variables are accounted for by the linear models in the leaves. We observe how the accuracy is drastically improved after including the linear models in the predictions.

\subsection{Power Management}
\label{SS:powermanagement}
Typically, the end customer receives a notification to curtail the power by some fraction. 
In this example on power management, we show how DPC can generate optimal inputs to track a desired power signal within a small allowance while maintaining the zone level thermal comfort. It may not be possible to have the same thermal comfort level in all the zones due to power curtailment, so we choose one zone (for example CEO's office) where the constraints must be met.
\begin{figure}[t!]
	\centering
	\includegraphics[width=30pc]{figures/eplus_validation.eps}	
	\caption{Model accuracy during training: The prediction made by forest using only $\X^d$ (red) captures the effect due to disturbances. The linear models in the leaves capture the local effects (green) due to the control inputs in $\X^c$ and improve the model accuracy.}
	\label{F:sepvars}
	\captionsetup{justification=centering}
\end{figure}
This is done by solving the following optimization problem, with control variables defined before:
\begin{align}
\begin{aligned}
& \underset{u_{k+j-1},\epsilon_j,\delta_j}{\mathrm{minimize}} & & \sum_{j=1}^{N} ({\tP}_{k+j} - \tP_{ref})^2 +  \lambda\epsilon_j + \nu \delta_j \\
& \mathrm{subject\ to }                                       & & \tP_{k+j} =  \hat{\Theta}_{\tP_j} [ 1,u^\top_{k},\ldots,u^\top_{k+j-1} ]^\top  \\
&                                                             & & \tT_{k+j} =  \hat{\Theta}_{\tT_j} [ 1,u^\top_{k},\ldots,u^\top_{k+j-1} ]^\top  \\
&                                                             & & \underline{\tP} - \epsilon_j \leq \tP_{k+j} \leq \bar{\tP} + \epsilon_j        \\
&                                                             & & \underline{\tT} - \delta_j   \leq \tT_{k+j} \leq \bar{\tT} + \delta_j          \\
&                                                             & & \underline{u}                \leq u_{k+j-1} \leq \bar{u}                       \\
&                                                             & & \epsilon_j \geq 0,\delta_j \geq 0, \ j = 1,\dots,N.
\end{aligned}\label{E:ptrack}
\end{align}

### 2018-01-20 paper/final/casestudyReal.tex
small changes

The comparison between numerical and experimental data is shown in Figure \ref{F:houseComparisonExperimental} and shows a quite good agreement. Therefore, with a MBE equal to $7.38\%$ and a CV(RMSE) equal to $8.37\%$, the EnergyPlus model of the use case can be considered well calibrated.  
\begin{figure}[t!]
	\begin{center}
		\subfigure[Thermal energy comparison. The accuracy error is $7.38\%$ with MBE definition and $8.37\%$ with CV(RMSE) definition]{
			\label{F:houseComparisonExperimentalEnergy}
			\centering
			\includegraphics[width=28pc]{figures/Thermal_energy_rev01.eps}
		}
		\subfigure[Thermal power comparison.]{
			\label{F:houseComparisonExperimentalPower}
			\centering
			\includegraphics[width=28pc]{figures/Thermal_power.eps}
		}
	\end{center}
	\caption{Comparison between numerical and experimental data.}
	\captionsetup{justification=centering}
	\label{F:houseComparisonExperimental}
\end{figure}
\subsection{Random forest models}
\label{SS:randomForestsModels}
For the closed-loop simulations with DPC we create $2$ different sets of models, $S_1 = \{\tT^{r_1}_{k+1},\tT^{r_2}_{k+1},\tT^{r_3}_{k+1},\tT^{r_4}_{k+j}\}$ and $S_2 = \{\tP_{k+j},\tT^{1}_{k+j},\tT^{2}_{k+j},\tT^{3}_{k+j},\\\tT^{4}_{k+j},\ j = 1,\ldots,N\}$, using random forests. In each set we have  $4$ models that describe the room temperature evolution ($\tT^{r_i}$ and $\tT^i$, $i=1,2,3,4$) in each of the $4$ rooms equipped with temperature sensors. In $S_2$ we have a model for the power consumption of the house ($\tP$). Models in $S_1$ are created using the classical random forests algorithm with all the features and are used as plant simulator of the house. For this reason, they are computed to give prediction only for time step $k+1$ and not over the whole horizon $N$. Models in $S_2$ are used as predictors over a horizon $N$ in the DPC algorithm and are then created using the methodology provided in Section \ref{SS:dpcrf}. The non-manipulated features in $\tX^d$ are the disturbance data (relative humidity, atmospheric pressure, outside air temperature, solar radiation, wind, time of the day and day of the week) and the states (temperature of the $4$ rooms). The manipulated feature in $\tX^c$ is the flow rate ($[m^3/h]$). All this features are used to create models in $S_1$ and the power model in $S_2$, while disturbance data, state temperature of room $i$ only and flow rate are used to identify $\hat{\Theta}^{T}_{\tT^i_j}$ for temperature models in $S_2$. All the models have been trained on the data from March $11$, $2016$ to April $26$, $2016$ and validated on the data from May $1$, $2016$ to May $15$, $2016$. The accuracy of these models with respect to real data is shown in Table \ref{T:S1accuracy}, based on the definition of Normalized Root Mean Square Error (NRMSE).

@@

\paragraph{DPC}
We want to optimize the ON/OFF heating system schedule in order to minimize power consumption of the house while keeping temperature of room $3$ within a comfort range. We also allow violations $\epsilon^{min},\epsilon^{max}$  of temperature bounds to guarantee feasibility of the algorithm. We include these violations in the objective function to be minimized.
\begin{table}[t!]
	\centering
	\begin{tabular}{cccccc}
		\toprule
		Set       & Power   & T. room $1$ & T. room $2$ & T. room $3$ & T. room $4$  \\ 
		\midrule
		$S_1$     &         & $96.58$     & $96.99$     & $97.21$     & $96.62$\\
		$S_2$     & $92.21$ & $97.38$     & $97.29$     & $96.93$     & $96.81$\\
		\bottomrule
	\end{tabular}
	\caption{Models accuracy for power and temperatures models in $S_1$ and $S_2$ expressed as $\mathrm{1-NRMSE}\,(\%)$.}
	\captionsetup{justification=centering}
	\label{T:S1accuracy}
	\vspace{-0.5cm}
\end{table}
\begin{figure}[t!]
	\begin{center}
		\subfigure[Power consumption model accuracy validation. The accuracy over the testing period expressed as $\mathrm{1-NRMSE}\,(\%)$ is $92.1\%$.]{
			\label{F:power_testing}
			\centering
			\includegraphics[width=26pc]{figures/power_testing.eps}
		}
		\subfigure[Temperature model accuracy validation for room $1$. The accuracy over the testing period expressed as $\mathrm{1-NRMSE}\,(\%)$ is $96.58\%$, for the temperature model $T^{r_1}$ in $S_1$ and $97.3\%$ for the temperature model $T^1$ in $S_2$. The accuracy for the other rooms is very similar as can be seen in Table \ref{T:S1accuracy}.]{
			\label{F:temperature_testing}
			\centering
			\includegraphics[width=26pc]{figures/temperature_testing.eps}
		}	
	\end{center}
	\vspace{-0.7cm}
	\caption{Model accuracy validation for power and temperature models.}
	\vspace{-0.7cm}
	\captionsetup{justification=centering}
	\label{F:testing}
\end{figure}

@@

\begin{problem}\label{P:dpcRealCase}
	\begin{align}
		\begin{aligned}
		& \underset{u_{k+j},\epsilon_j^{min},\epsilon_j^{max}}{\text{minimize}} & & \sum_{j=1}^{N} Q{\tP}_{k+j}^2 + \lambda_{min}\parallel\epsilon^{min}\parallel_2 + \lambda_{max}\parallel\epsilon^{max}\parallel_2                                                                                                                \\
		& \text{subject to }                             & & \tP_{k+j}   = \hat{\Theta}_{\tP_j}   [ 1,u_k,\ldots,u_{k+j-1} ]^\top                                      \\
		&                                                & & \tT^1_{k+j} = \hat{\Theta}_{\tT^1_j} [ 1,u_k,\ldots,u_{k+j-1} ]^\top                                      \\
		&                                                & & \tT^2_{k+j} = \hat{\Theta}_{\tT^2_j} [ 1,u_k,\ldots,u_{k+j-1} ]^\top                                      \\
		&                                                & & \tT^3_{k+j} = \hat{\Theta}_{\tT^3_j} [ 1,u_k,\ldots,u_{k+j-1} ]^\top                                      \\
		&                                                & & \tT^4_{k+j} = \hat{\Theta}_{\tT^4_j} [ 1,u_k,\ldots,u_{k+j-1} ]^\top                                      \\
		&                                                & & \underline{\tT}_{k+j-1}-\epsilon^{min}_j \leq \tT^3_{k+j} \leq \overline{\tT}_{k+j-1} + \epsilon^{max}_j  \\
		&                                                & & u_{k+j-1}   = \underline{u} \lor u_{k+j-1} = \overline{u}                                                 \\
		&                                                & & \epsilon^{min}_j,\epsilon^{max}_j \geq 0, \ j = 1,\dots,N.
		\end{aligned}
	\end{align}\label{E:DPCrealcase}
\end{problem}

>>>>

The comparison between numerical and experimental data is shown in Figure \ref{F:houseComparisonExperimental} and shows a quite good agreement. Therefore, with a MBE equal to $7.38\%$ and a CV(RMSE) equal to $8.37\%$, the EnergyPlus model of the use case can be considered well calibrated.  
\begin{figure}[t!]
	\begin{center}
		\subfigure[Thermal energy comparison. The accuracy error is $7.38\%$ with MBE definition and $8.37\%$ with CV(RMSE) definition]{
			\label{F:houseComparisonExperimentalEnergy}
			\centering
			\includegraphics[width=28pc]{figures/Thermal_energy_rev01.eps}
		}
		\subfigure[Thermal power comparison.]{
			\label{F:houseComparisonExperimentalPower}
			\centering
			\includegraphics[width=28pc]{figures/Thermal_power.eps}
		}
	\end{center}
	\caption{Comparison between numerical and experimental data.}
	\captionsetup{justification=centering}
	\label{F:houseComparisonExperimental}
\end{figure}
\subsection{Random forest models}
\label{SS:randomForestsModels}
For the closed-loop simulations with DPC we create $2$ different sets of models, $S_1 = \{\tT^{r_1}_{k+j},\tT^{r_2}_{k+j},\tT^{r_3}_{k+j},\tT^{r_4}_{k+j}\}$ and $S_2 = \{\tP_{k+j},\tT^{1}_{k+j},\tT^{2}_{k+j},\tT^{3}_{k+j},\tT^{4}_{k+j},\ j = 1,\ldots,N\}$, using random forests. In each set we have  $4$ models that describe the room temperature evolution ($\tT^{r_i}$ and $\tT^i$, $i=1,2,3,4$) in each of the $4$ rooms equipped with temperature sensors. In $S_2$ we have a model for the power consumption of the house ($\tP$). Models in $S_1$ are created using the classical random forests algorithm with all the features and are used as plant simulator of the house. For this reason, they are computed to give prediction only for time step $k+1$ and not over the whole horizon $N$. Models in $S_2$ are used as predictors over a horizon $N$ in the DPC algorithm and are then created using the methodology provided in Section \ref{SS:dpcrf}. The non-manipulated features in $\X^d$ are the disturbance data (relative humidity, atmospheric pressure, outside air temperature, solar radiation, wind, time of the day and day of the week) and the states (temperature of the $4$ rooms). The manipulated feature in $\X^c$ is the flow rate ($[m^3/h]$). All this features are used to create models in $S_1$ and the power model in $S_2$, while disturbance data, state temperature of room $i$ only and flow rate are used to identify $\hat{\Theta}_{\tT^i_j}$ for temperature models in $S_2$. All the models have been trained on the data from March $11$, $2016$ to April $26$, $2016$ and validated on the data from May $1$, $2016$ to May $15$, $2016$. The accuracy of these models with respect to real data is shown in Table \ref{T:S1accuracy}, based on the definition of Normalized Root Mean Square Error (NRMSE).

@@

\paragraph{DPC}
We want to optimize the ON/OFF heating system schedule in order to minimize power consumption of the house while keeping temperature of room $3$ within a comfort range. We also allow violations $\epsilon^{min},\epsilon^{max}$  of temperature bounds to guarantee feasibility of the algorithm. We include these violations in the objective function to be minimized.
\begin{table}[t!]
	\centering
	\begin{tabular}{cccccc}
		\toprule
		Set       & Power   & T. room $1$ & T. room $2$ & T. room $3$ & T. room $4$  \\ 
		\midrule
		$S_1$     &         & $96.58$     & $96.99$     & $97.21$     & $96.62$\\
		$S_2$     & $92.21$ & $97.38$     & $97.29$     & $96.93$     & $96.81$\\
		\bottomrule
	\end{tabular}
	\caption{Models accuracy for power and temperatures models in $S_1$ and $S_2$ expressed as $\mathrm{1-NRMSE}\,(\%)$.}
	\captionsetup{justification=centering}
	\label{T:S1accuracy}
	\vspace{-0.5cm}
\end{table}

@@

\begin{problem}\label{P:dpcRealCase}
	\begin{align}
		\begin{aligned}
		& \underset{u_{k+j-1},\epsilon_j^{min},\epsilon_j^{max}}{\mathrm{minimize}} & & \sum_{j=1}^{N} Q{\tP}_{k+j}^2 + \lambda_{min}\parallel\epsilon^{min}\parallel_2 + \lambda_{max}\parallel\epsilon^{max}\parallel_2                            \\
		& \mathrm{subject\ to } & & \tP_{k+j}   = \hat{\Theta}_{\tP_j}   [ 1,u_k,\ldots,u_{k+j-1} ]^\top                         \\
		&                    & & \tT^1_{k+j} = \hat{\Theta}_{\tT^1_j} [ 1,u_k,\ldots,u_{k+j-1} ]^\top                            \\
		&                    & & \tT^2_{k+j} = \hat{\Theta}_{\tT^2_j} [ 1,u_k,\ldots,u_{k+j-1} ]^\top                            \\
		&                    & & \tT^3_{k+j} = \hat{\Theta}_{\tT^3_j} [ 1,u_k,\ldots,u_{k+j-1} ]^\top                            \\
		&                    & & \tT^4_{k+j} = \hat{\Theta}_{\tT^4_j} [ 1,u_k,\ldots,u_{k+j-1} ]^\top                            \\
		&                    & & \underline{\tT}_{k+j}-\epsilon^{min}_j\leq\tT^3_{k+j}\leq\overline{\tT}_{k+j}+\epsilon^{max}_j  \\
		&                    & & u_{k+j-1}   = \underline{u} \lor u_{k+j-1} = \overline{u}                                       \\
		&                    & & \epsilon^{min}_j,\epsilon^{max}_j \geq 0, \ j = 1,\dots,N.
		\end{aligned}
	\end{align}\label{E:DPCrealcase}
\end{problem}

\begin{figure}[t!]
	\begin{center}
		\includegraphics[width=26pc]{figures/power_testing.eps}
		\caption{Power consumption model accuracy validation. The accuracy over the testing period expressed as $\mathrm{1-NRMSE}\,(\%)$ is $92.1\%$.}
		\label{F:power_testing}			
	\end{center}
\end{figure}

\begin{figure}[t!]
	\begin{center}
		\includegraphics[width=26pc]{figures/temperature_testing.eps}
		\caption{Temperature model accuracy validation for room $1$. The accuracy over the testing period expressed as $\mathrm{1-NRMSE}\,(\%)$ is $96.58\%$, for the temperature model $T^{r_1}$ in $S_1$ and $97.3\%$ for the temperature model $T^1$ in $S_2$. The accuracy for the other rooms is very similar as can be seen in Table \ref{T:S1accuracy}.}
		\label{F:temperature_testing}	
	\end{center}
\end{figure}

### 2018-01-20 paper/final/cs-comparison.tex
small changes

\begin{table}[h!]
	\centering
	\begin{tabular}{ccccc}
		\toprule
		& explained & mean objective& mean input  & mean  \\
		&  variance$[\mathrm{-}]$ & value $[\mathrm{-}]$ & cost $[-]$ & deviance $[\mathrm{^oC}]$ \\     
		\midrule
		MPC    &  $\mathrm{-}$ &  22.60 & 17.16  &  0.26  \\
		DPC-En   & 70.1\% &  39.26  & 15.12 &  0.48 \\
		DPC-RT  & 1.8\% & 204.55 & 16.84 &  0.57 \\
		\bottomrule
	\end{tabular}
	\vspace{0.2cm}
	\caption{Quantitative comparison of explained variance, mean value of objective function, mean input cost $c^Tu$ and mean deviance from the reference temperature $|\mathsf{T}-\mathsf{T}_{\mathrm{ref}}|$.}
	\captionsetup{justification=centering}
	\label{T:comparison}
\end{table}

\begin{figure}[t!]
	\begin{center}
	\vspace{1.1cm}
	\subfigure[External disturbances: solar gain, internal gain due to equipment and dry-bulb temperature.]{
		\label{F:dist}
		\centering
		\includegraphics[width=25pc]{figures/disturbances.eps}
	}

	\subfigure[Optimal control input: cooling usage factor $\mathsf{C}$ with  $0 \leq \mathsf{C} \leq 1$. DPC-En generates a control strategy very simular to MPC.]{
		\label{F:control1}
		\centering
		\includegraphics[width=25pc]{figures/input3.eps}
	}

\end{center}
\end{figure}
\begin{figure}[h!]
\begin{center}
	\subfigure[Optimal control input: radiator heat $\mathsf{H}$ with  $0 \leq \mathsf{H} \leq 23 \ \mathrm{W/m^2}$. Again, DPC-En generates a control strategy very simular to MPC.]{
		\label{F:control2}
		\includegraphics[width=25pc]{figures/input4.eps}
	}

	\subfigure[Room temperature has time varying bounds. When the building is occuped the constraints are relaxed, else $19(20) \leq \mathsf{T}_{\mathrm{in}} \leq 25(26) \ \mathrm{^oC}$ in January(May). MPC and DPC-En are able to track the reference temperature ($22 \ \mathrm{^oC}$) closely.]{
		\label{F:state}
		\centering
		\includegraphics[width=25pc]{figures/state.eps}
	}

	\subfigure[Cumulative optimal cost after solving optimization. MPC serves as the benchmark with the minimum cost, followed by DPC-En and then DPC-RT.]{
		\label{F:obj}
		\centering
		\includegraphics[width=25pc]{figures/cumsumcost.eps}
	}
	\end{center}
	\vspace{-0.6cm}
	\caption{Comparison of optimal performance obtained with MPC, DPC-En and DPC-RT for 3 days in January and 3 days in May.}
	\label{F:comparison}
	\captionsetup{justification=centering}
	\vspace{-0.3cm}
\end{figure}
Thus, we have shown that DPC-En provides a comparable performance to MPC without using the physical model.
However, one major limitation of the bilinear model is that the information about the building power consumption is not available. Much nonlinearities in the system are due to equipment efficiencies which are not considered in the bilinear case but are very important for practical purposes.

>>>>

\begin{table}[h!]
	\centering
	\begin{tabular}{ccccc}
		\toprule
		& explained & mean objective& mean input  & mean  \\
		&  variance$[\mathrm{-}]$ & value $[\mathrm{-}]$ & cost $[-]$ & deviance $[\mathrm{^oC}]$ \\     
		\midrule
		MPC    &  $\mathrm{-}$ &  22.60 & 17.16  &  0.26  \\
		DPC-En   & 70.1\% &  39.26  & 15.12 &  0.48 \\
		DPC-RT  & 1.8\% & 204.55 & 16.84 &  0.57 \\
		\bottomrule
	\end{tabular}
	\vspace{0.2cm}
	\caption{Quantitative comparison of explained variance, mean value of objective function, mean input cost $c^\top u$ and mean deviance from the reference temperature $|\mathsf{T}-\mathsf{T}_{\mathrm{ref}}|$.}
	\captionsetup{justification=centering}
	\label{T:comparison}
\end{table}

\begin{figure}[t!]
	\begin{center}
	\vspace{1.1cm}
	\includegraphics[width=23pc]{figures/disturbances.eps}
	\caption{External disturbances: solar gain, internal gain due to equipment and dry-bulb temperature.}
	\label{F:dist}
\end{center}
\end{figure}
\begin{figure}[h!]
\begin{center}
		\subfigure[Optimal cooling factor $\mathsf{C}$. DPC-En control strategy very similar to MPC.]{
		\label{F:control1}
		\centering
		\includegraphics[width=23pc]{figures/input3.eps}
	}

	\subfigure[Optimal radiator heat $\mathsf{H}$. DPC-En control strategy very similar to MPC.]{
		\label{F:control2}
		\includegraphics[width=23pc]{figures/input4.eps}
	}

	\subfigure[Room temp. has time varying bounds. When building is occupied, constraints are relaxed. MPC and DPC-En track the ref. temp. ($22 \ \mathrm{^oC}$) closely.]{
		\label{F:state}
		\centering
		\includegraphics[width=23pc]{figures/state.eps}
	}

	\subfigure[Cumulative optimal cost after solving optimization. MPC serves as the benchmark with the minimum cost, followed by DPC-En and then DPC-RT.]{
		\label{F:obj}
		\centering
		\includegraphics[width=23pc]{figures/cumsumcost.eps}
	}
	\end{center}
	\vspace{-0.6cm}
	\caption{Comparison of optimal performance obtained with MPC, DPC-En and DPC-RT for 3 days in January and 3 days in May.}
	\label{F:comparison}
	\captionsetup{justification=centering}
	\vspace{-0.3cm}
\end{figure}
Thus, we have shown that DPC-En provides a comparable performance to MPC without using the physical model.
However, one major limitation of the bilinear model is that the information about the building power consumption is not available. Much nonlinearities in the system are due to equipment efficiencies which are not considered in the bilinear case but are very important for practical purposes.
